---
title: "ADA Final Project"
author: "Tori"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in the data and necessary packages

```{r}
pacman::p_load(haven, dplyr, tidyverse, ggplot2, ggfortify, broom, odds.n.ends, blorr, nnet, car, table1, caret, ggcorrplot, gtsummary, expss)

#(This step from the orignal full brfss 2019 file)
#brfss19 <- read_xpt("LLCP2019.xpt")
#brfss19

brfss19cut <-read_csv("brfss19cut.csv")
brfss19cut
```

## Management(from original full dataset found here: https://www.cdc.gov/brfss/annual_data/annual_2019.html)

```{r}
###This was management from the original brfss 2019 dataset- I made it into a smaller csv file below so these steps are not needed
#Getting rid of underscores in front of variable name
#names(brfss19) <- gsub("\\_", "", names(brfss19))

#Recoding variables for age, sex, race, CHD status, BMI, exercise frequency, and general health. Also recoded and categorized formal exercise variable. 
#brfss19 <- brfss19 %>%
  #mutate(age = as.numeric(AGE80),
#         sex = factor(SEXVAR, levels = c(1,2), labels = c("Male", "Female")),
#         race = case_when(IMPRACE %in% 1 ~ "White",
#                          IMPRACE %in% 2 ~ "Black",
#                          IMPRACE %in% 3 ~ "Asian",
#                          IMPRACE %in% 4 ~ "American Indian/Alaskan Native",
#                          IMPRACE %in% 5 ~ "Hispanic",
#                          IMPRACE %in% 6 ~ "Other"),
#         race = factor(race, levels = c("White", "Black", "Asian", "American Indian/Alaskan Native", "Hispanic", "Other")),
#         BMI = as.numeric(BMI5),
#         ex_freq = as.numeric(PAFREQ1),
#         gen_health = case_when(GENHLTH %in% 1 ~ "Excellent",
#                                GENHLTH %in% 2 ~ "Very good",
#                                GENHLTH %in% 3 ~ "Good",
#                                GENHLTH %in% 4 ~ "Fair",
#                                GENHLTH %in% 5 ~ "Poor"),
#         gen_health = factor(gen_health, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
#         ex_formal = case_when(EXRACT11 %in% c(7, 37) ~ "Yes",
#                               EXRACT11 %in% c(18, 71, 73, 76) ~ "No"),
#         ex_formal = factor(ex_formal, levels = c("No", "Yes")),
#         ever_smk = case_when(SMOKER3 %in% c(1, 3) ~ 1,
#                              SMOKER3 %in% c(2, 4) ~ 2),
#         ever_smk = factor(ever_smk, levels = c(1, 2), labels = c("Yes", "No")),
#        )

#Per the codebook, the exercie frequency variable has 3 implied decimal places. I want to insert those decimal places explicitly:
#brfss19$ex_freq <- brfss19$ex_freq/1000

#Selecting just the variables needed and putting them in a new dataset
#brfss19cut <- brfss19 %>%
#  filter(!STATE %in% c(66, 72)) %>%
#  select(age, sex, race, CHD, ex_formal, ever_smk, ex_freq) %>%
#  na.omit()

#Changing the reference level of CHD
#brfss19cut$CHD <- relevel(brfss19cut$CHD, ref = "No")
#brfss19cut$ever_smk <- relevel(brfss19cut$ever_smk, ref = "No")
#brfss19cut
#summary(brfss19cut)

#Making a smaller dataset to work with
#write.csv(brfss19cut,"brfss19cut.csv")
```

## Checking the recoded variables
```{r}
table(brfss19cut$ex_formal)
table(brfss19cut$race)
table(brfss19cut$ever_smk)
```

##Table 1
```{r}
#Adding variable labels
label(brfss19cut$ex_formal) <- "Formal exercise"
label(brfss19cut$age) <- "Age(yrs)"
label(brfss19cut$sex) <- "Sex"
label(brfss19cut$race) <- "Race"
label(brfss19cut$ex_freq) <- "Exercise frequency(per week)"
label(brfss19cut$ever_smk) <- "Ever smoker"
label(brfss19cut$CHD) <- "Coronary Heart Disease(CHD)"

#Creating table 1
table1(~ex_formal + age + sex + race + ex_freq + ever_smk | CHD, brfss19cut, render.missing = NULL)
```

## Binary/Multivariate Logistic Regression analysis of CHD status and exercise type 
```{r}
#Logistic model for CHD (before adjusting)
model_CHD_before <- glm(CHD ~ ex_formal, family = binomial, data = brfss19cut)
model_CHD_before
#exponentiating model coefficients for odds
exp(coef(model_CHD_before))

#Logistic model for CHD (after adjusting)
model_CHD <- glm(CHD ~ ex_formal + age + sex + race + ex_freq + ever_smk, family = binomial, data = brfss19cut)
model_CHD

#The odds are very different before vs. after adjusting for covariates

#Calculate and print the Odds Ratios and 95%CIs
odds.n.ends(model_CHD)
```

## Testing for Effect modification
```{r}
model_em <- glm(CHD ~ ex_formal + age + sex + race + ex_freq + ever_smk + sex*ex_formal, family = binomial, data = brfss19cut)
model_em

OR_em<-exp(cbind(OR = coef(model_em), confint(model_em)))
summary(OR_em)
odds.n.ends(model_em)

model_em2 <- glm(CHD ~ ex_formal + age + sex + race + ex_freq + ever_smk*ex_formal, family = binomial, data = brfss19cut)
model_em2

OR_em2<-exp(cbind(OR = coef(model_em2), confint(model_em2)))
summary(OR_em2)
odds.n.ends(model_em2)

#Interaction terms were not significant- no effect modification by sex or smoking for exercise frequency and CHD. 
```

## Testing assumptions. 
```{r}

#Correlations
OHE <- dummyVars(" ~ .", data = brfss19cut,sep = "__")
chd2 <- data.frame(predict(OHE,newdata = brfss19cut))
chd2
corMat <- cor(chd2)
ggcorrplot(corMat)
#Model VIFs
vif(model_CHD)
#All variables have VIF values well below 5- no multicollinearity
```

```{r}
#Linearity
#Testing linearity of age and CHD
brfss19cut <- brfss19cut %>%
  mutate(age.times.logage = age * log(age)) #create term to test linearity

boxTidwellage <- glm(CHD ~ age + age.times.logage, data = brfss19cut, family = "binomial") #Box Tidwell technique, test the assumption of linearity

summary(boxTidwellage)
#Testing linearity of exercise frequency and CHD
brfss19cut <- brfss19cut %>%
  mutate(ex_freq.times.logex_freq = ex_freq * log(ex_freq)) #create term to test linearity

boxTidwellex_freq <- glm(CHD ~ ex_freq + ex_freq.times.logex_freq, data = brfss19cut, family = "binomial") #Box Tidwell technique, test the assumption of linearity

summary(boxTidwellex_freq)
```

```{r}
#Influence plot- testing for disproportionate influence
plot(model_CHD, which=4, id.n=5, col="red")
#Cook's D cutoff (4/n)
cutoff <- 0.00015
#identify observations with a Cook's Distance greater than cutoff
obs_no <- as.data.frame(cooks.distance(model_CHD)) %>%
  mutate(obs_no=row_number()) %>%
  filter(`cooks.distance(model_CHD)` > cutoff)
#Excluding values in Cook's D plot for comparison
model_CHD.modex <- update(model_CHD,subset=c(-obs_no$obs_no))
summary(model_CHD.modex)
#Compare coefficients between models with and without influential observations

compareCoefs(model_CHD, model_CHD.modex) 
  #Removing influential data largely affects the coefficients, however, all the influential points are cases and important for analyses
```

```{r}
#Model fits 
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(model_CHD)

#Hosmer Lemeshow goodness of fit
blr_test_hosmer_lemeshow(model_CHD)
```
  ### Interpretation: Based on the model results, the odds of CHD is 38% lower in those that mainly participate in formal exercise than those that mainly participate in informal exercise. 


## Formatted table of regression model results
```{r}

tbl_regression(model_CHD, exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2), 
               label = list(ex_formal ~ "Formal exercise",
                 age ~ "Age(yrs)",
                 sex ~ "Sex",
                 race ~ "Race",
                 ex_freq ~ "Exercise frequency(per week)",
                 ever_smk ~ "Ever smoker"
                            ),
  ) %>% 
  bold_labels() %>%
  italicize_levels()
```





